Pathogen Spillover - dissertation chapter

### 2019-04-07 pathogen-spillover-diss-ch.Rmd
Old update?

---
title: "Pathogen Spillover - dissertation chapter"
author: "Whalen Dillon"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: no
    toc: true
    toc_float: true
  pdf_document: default
  word_document: default
---

@@

![Conceptual Path Model](figures/Fig2-conceptual-path-model-updated.png)

@@

```{r load data, echo=F, message=F}
library(knitr)
load("climate_path_model_data.RData")

# devtools::install_github("jslefche/piecewiseSEM@devel", build_vignette = TRUE)
library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
## Avoided this error in 1.1.2 by removing observations with NA values, which gives the same results as before, so maybe it wasn't properly dropping missing values when extracting info from the models.
library(plyr); library(dplyr, verbose = F)
library(lme4, verbose = F); library(nlme, verbose = F)
library(ggplot2)
library(viridisLite); library(viridisLite)
# library(lmerTest, verbose = F)

@@

## make consolidated data frame only of variables in the path model
oakplots.model.data <- select(oakplots.sub, plotid, sample_year, oak.inf, inf_oak_ct, uninf_oak_ct, tot_bay, umca_density_ha, tot_lfct, tot_lfct.log, hrs_abv25ds, avg_hrs1422_wet_t_t1, avg_hrs1422_wet_t_t1.log, umca_density_ha.log, H.2005, twi15m, twi15m.log#, twi15m.log.scl, tot_lfct.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, umca_density_ha.log.scl, shannons2005.scl
                              )

oakplots.model.data <- filter(oakplots.model.data, avg_hrs1422_wet_t_t1 !="NA", H.2005 != "NA", tot_lfct != "NA")
write.csv(oakplots.model.data, "analysis/data/oakplots-pathmodel-data.csv",
          row.names = F)

# Exclude LIDE plots
oakplots.model.data.sub <- oakplots.model.data %>% 
      filter(plotid!="hood03", plotid!="hood05", plotid!="jlsp04", plotid!="johns02", plotid!="sugar17")

# write.csv(oakplots.model.data.sub, "pathogen_spillover_data.csv", row.names = F)

```

# Path modeling 
This is the list of models that make up the path model to be fit using the methods implemented in the 'piecewiseSEM' packag. The continuous variables in the models were standardized using the default settings of the 'scale' function in R, resulting in each having a mean of zero and standard deviation of one. Each variable is first centered by subtracting its mean and then scaled by dividing by its standard deviation.

@@

```{r hours >25 path model all oak plots unscaled, echo=F, eval=F}
# slc_nlme <- lme(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005, random = ~1|plotid, data = oakplots.model.data)

model_list1 <- list(
      shannons <- lm(
            H.2005 ~ twi15m, 
            data = filter(oakplots.model.data, sample_year==2006)),

      umca_density <- lmer(
            umca_density_ha.log ~ twi15m + (1|plotid) + (1|sample_year), 
            data = oakplots.model.data),

@@

      wet_hours_1422 <- lmer(
            avg_hrs1422_wet_t_t1.log ~ twi15m + (1|plotid) + (1|sample_year), 
            data = oakplots.model.data),

      hrs_abv25 <- lmer(
            hrs_abv25ds ~ twi15m + (1|plotid) + (1|sample_year), 
            data = oakplots.model.data),

      oak_infection <- glmer(
            oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), 
            data = oakplots.model.data)
)
```

Add some re-calcualted variables and explore the bay density~diversity relationship.

```{r revise data names}
oaks.data <- oakplots.model.data %>% 
      rename(hrs25 = hrs_abv25ds, 
             hrs1422 = avg_hrs1422_wet_t_t1.log, 
             lfct = tot_lfct.log, 
             umcadens = umca_density_ha.log, 
             h2005 = H.2005, 
             oakinf = oak.inf, 
             year = sample_year)

oaks.data <- oaks.data %>% 
      mutate(tot_bay.ln = log(tot_bay),
             tot_bay.ln1 = log1p(tot_bay),
             tot_bay0.1 = ifelse(tot_bay==0, 0.1, tot_bay),
             tot_bay.ln.1 = log(tot_bay0.1)
             )
```

```{r explore diveristy density}

## refit model with missing path from diversity to umca density

library(GGally)
ggpairs(select(oaks.data, h2005, twi15m, tot_bay, tot_bay.ln, tot_bay.ln1))

ggplot(oaks.data, 
       aes(h2005, tot_bay)) +
      geom_point() +
      geom_smooth(method = "lm") +
      scale_fill_viridis_d() +
      scale_color_viridis_d() +
      # geom_density(alpha = .8) +
      theme_classic() +
      NULL

ggplot(oaks.data, aes(tot_bay)) +
      geom_density() +
      theme_classic()
ggplot(oaks.data, aes(tot_bay.ln)) +
      geom_density() +
      geom_density(aes(tot_bay.ln1), color = "blue") +
      theme_classic()

ggplot(oaks.data) +
      geom_density(aes(h2005)) +
      theme_classic()
```

```{r explore models, eval=F}

oaks.data %>% 
      # filter(tot_bay>0) %>% 
      filter(year==2006) %>%
ggplot(aes(h2005, tot_bay.ln)) +
      geom_point() +
      geom_smooth(method = "lm") +
      geom_smooth(aes(y=tot_bay.ln1), method = "lm", color = "red") +
      # geom_smooth(aes(y=tot_bay.ln.001), method = "lm", color = "tan4") +
      # geom_density(alpha = .8) +
      theme_classic() +
      NULL
## DON'T ADD 1 & LOG-TRANSFORM
summary(oaks.data)
filter(oaks.data, is.infinite(tot_bay.ln))$tot_bay
oaks.data$tot_bay.ln[is.infinite(oaks.data$tot_bay.ln)] <- 0

m1 <- lmer(tot_bay ~ h2005 + twi15m + (1|plotid) + (1|year),
           data = oaks.data, na.action = na.omit)
m1a <- lmer(tot_bay.ln ~ h2005 + twi15m + (1|plotid) + (1|year),
           data = oaks.data, na.action = na.omit)
# anova(m1,m1a)
summary(m1)
plot(m1)
# plot(m1a)
# summary(m1a)

m2 <- mgcv::gamm(tot_bay ~ s(h2005) + twi15m, random=list(year=~1, plotid=~1), data = oaks.data, na.action = na.omit)
summary(m1)
summary(m2[[1]])
summary(m2[[2]])
plot(m2[[1]])
plot(m2[[2]])
mgcv::vis.gam(m2)
m3 <- gamm4::gamm4(tot_bay ~ s(h2005) + twi15m, random = ~(1|plotid)+(1|year), data = oaks.data, na.action = na.omit)
summary(m3[[1]])
plot(m3[[1]])
summary(m3[[2]])
plot(m3[[2]])

m4 <- glmer(tot_bay ~ h2005 + twi15m + (1|plotid) + (1|year),
           data = oaks.data, na.action = na.omit, family = poisson)
# par(mfrow=c(2,1))
plot(m1); plot(m4)
summary(m1)
summary(m4)
anova(m1, m4)

m5 <- glmer.nb(tot_bay ~ h2005 + twi15m + (1|plotid) + (1|year),
           data = oaks.data, na.action = na.omit)
summary(m5)
plot(m5)

dat <- filter(oaks.data, year==2006)
m6 <- glm(tot_bay ~ h2005 + twi15m, data = dat, family = quasipoisson, na.action = na.fail)
summary(m6)
# plot(m6)
(c_hat <- deviance(m6) / df.residual(m6))

library(MuMIn)
dredge(m6, rank = "QAIC", chat = c_hat)
dredge(m6, rank = "AIC")

hacked.quasipoisson <- function(...) {
    res <- quasipoisson(...)
    res$aic <- poisson(...)$aic
    res
}

QAIC(update(m6, family = hacked.quasipoisson, chat = c_hat)

# library(gamlss)
# m7 <- gamlss(tot_bay ~ h2005 + twi15m, data = oaks.data)
# summary(m7)
# plot(m7)

# GGally::ggpairs(select(oaks.data, tot_lfct,tot_bay,h2005))

m1 <- glmer(tot_lfct ~ tot_bay + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), family = poisson, data = oaks.data, na.action = na.omit)
summary(m1)

m2 <- glmer.nb(tot_lfct ~ tot_bay + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit)

summary(m2)
anova(m2,lfctmod)
plot(lfctmod)

```

```{r refit model with H bay density path, eval=F}
# The refit model has path between H and UMCA density ####
model_list1_refit <- list(
      shannonmod <- lm(h2005 ~ twi15m, data = filter(oaks.data, year==2006), na.action = na.omit),

      densmod <- lmer(umcadens ~ h2005 + twi15m + (1|plotid) + (1|year),
           data = oaks.data, na.action = na.omit),

      lfctmod <- lmer(lfct ~ umcadens + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),

      wetmod <- lmer(hrs1422 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),

      hotmod <- lmer(hrs25 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),

      # hrs25 %~~% hrs1422,

      infmod <- glmer(oakinf ~ lfct + hrs1422 + h2005 + (1|plotid) + (1|year), family = binomial(link = "logit"), data = oaks.data, na.action = na.omit)
)
summary(lmer(h2005 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit))
summary(densmod)
summary(lfctmod)
summary(wetmod)
summary(hotmod)
summary(infmod)

# summary(oaks.data)

# ?psem I get errors with row matching when trying to use the new fucntion. I think it may have more to do with the new version of R and how that affects the internal sub-modeling when there are complex mixed effects models.

```

```{r Update model list}

# Updated SEM uses untransformed UMCA variables ####
# lfct is the ln(tot_lfct+1)
      # lmer(h2005 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
m_list_update <- list(

            shannon.mod <- lm(h2005 ~ twi15m, data = filter(oaks.data, year==2006), na.action = na.omit),
      # changed the bay response to be log-transformed 
      umca.mod <- lmer(tot_bay.ln1 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),

      lfct.mod <- lmer(lfct ~ tot_bay.ln1 + hrs25 + hrs1422 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),

      wet.mod <- lmer(hrs1422 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),

      hot.mod <- lmer(hrs25 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),

      # hrs25 %~~% hrs1422,

      inf.mod <- glmer(oakinf ~ lfct + hrs1422 + h2005 + (1|plotid) + (1|year), family = binomial(link = "logit"), data = oaks.data, na.action = na.omit)

)

# ggplot(oaks.data, aes(tot_bay.ln1, lfct)) +
#       geom_point() +
#       geom_smooth(method = "lm", color = "red") +
#       geom_smooth() +
#       theme_classic()

```

```{r path-model-no-lide-plots, echo=F, eval=F}
model_list1.noLIDE <- list(
      shannons <- lm(H.2005 ~ twi15m, data = filter(oakplots.model.data.sub, sample_year == 2006)),

      umca_density <- lmer(umca_density_ha.log ~ twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),

      slc <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ 
                                   # umca_density_ha.log + twi15m.log +
                                   twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),

      hrs_abv25 <- lmer(hrs_abv25ds ~ 
                              # umca_density_ha.log + 
                              twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data.sub)
)

```

@@

# Examining the residual plots of each piece of the path model

## **Residual plots of Shannon's Diversity Index model**

### Diagnostic plots 
```{r shannon.mod diagnostic plots, echo=F}

library(sjPlot)
plot_model(shannon.mod, type = "diag", show.values = T, grid = TRUE)
plot_model(shannon.mod, type = "resid", show.values = T)

```

### Fitted model plots
```{r shannon.mod fit plots}

plot_model(shannon.mod, type = "slope", show.values = T)
plot_model(shannon.mod, type = "std", show.values = T, 
           title = "Standardized coefficient estimate")
S.plot <- plot_model(shannon.mod, type = "pred", terms = "twi15m")

# line_color <- "black"
# S.plot$data
# str(S.plot)

(shannon.plot <- plot_model(shannon.mod, type = "pred", terms = "twi15m") +
      geom_point(data = oaks.data, 
                 aes(twi15m, h2005), 
                 alpha = .1, size = 1, position = "jitter") +
      # geom_smooth(method = "loess", se = F, color = "gray50", linetype = "dashed") +
      # geom_smooth(method = "lm", color = "black") +
      theme_classic() +
      xlab("Topographic index") +
      ylab("Tree diversity (H')") +
      theme(strip.background = element_blank(),
            strip.text = element_blank()) +
      NULL
      )

# plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
# plot(shannons$model$twi15m, resid(shannons),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "Diversity model residuals vs. wetness index predictor")

```

## **Residual plots of bay laurel density model**

### Diagnostic plots
```{r umca.mod diagnostic plots, echo=F}

plot_model(umca.mod, type = "diag")
plot_model(umca.mod, type = "resid")

```

### Fitted model plots
```{r umca.mod fit plots}

plot_model(umca.mod, show.values = T, type = "est", 
           title = "Coefficient estimates - UMCA density")
plot_model(umca.mod, show.values = T, type = "std", 
           title = "Standardized coefficient estimates - UMCA density")
plot_model(umca.mod, show.values = T, type = "re")

# umca_vs_h2005 <- plot_model(umca.mod, show.values = T, type = "pred", terms = c("twi15m"), pred.type = "fe")
# umca_vs_h2005 +
#       theme_classic()

# oaks.data %>% 
#       filter() %>% 
#       ggplot(aes(h2005, log1p(tot_bay))) +
#       geom_point() +
#       geom_smooth(method = "lm") +
      # facet_grid(.~year)
# qplot(x=h2005, y=tot_bay, data = oaks.data)

# library(ggeffects)
# umca.mod_predict <- ggpredict(model = umca.mod.update, terms = "h2005", type = "re", full.data = T)
# ggplot(umca.mod_predict, aes(x, predicted)) +
#       # data = umca.mod.update@frame,
#       geom_point(aes(x, predicted), show.legend = F) +
#       # geom_smooth(se = F) +
#       geom_smooth(method = "lm") +
#       # facet_grid(.~year) +
#       xlab("Shannon's diversity index (H')") +
#       ylab("Bay laurel density (stems/ha)") +
#       theme_classic() +
#       NULL

# oaks.data$h2005.sq <- oaks.data$h2005^2
# oaks.data$tot_bay.log <- log1p(oaks.data$tot_bay)
# umca.mod.update <- lmer(tot_bay.log ~ h2005 + twi15m + (1 | plotid) + (1 | year), data = oaks.data)
# plot(umca.mod)
# plot_model(umca.mod.update, show.values = T, type = "resid")
# plot_model(umca.mod.update, show.values = T, type = "diag")

# umca_vs_twi15m <- plot_model(densmod, show.values = T, type = "eff", terms = c("twi15m"), pred.type = "re")

# plot(umca_density, main = "LMM of bay laurel density residuals")
# plot(umca_density@frame$twi15m, resid(umca_density),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "Bay laurel density model residuals vs. wetness index predictor")

```

## **Residual plots of symptomatic leaf count model**

### Diagnostic plots
```{r lfct.mod plot, echo=F}

# diagnostic plots
plot_model(lfct.mod, type = "diag")
plot_model(lfct.mod, type = "resid")

```

### Fitted model plots
```{r lfct.mod fit plots}
# estimate plots
plot_model(lfct.mod, show.values = T, type = "est", 
           title = "Coefficient estimates - leaf count")
plot_model(lfct.mod, show.values = T, type = "std", 
           title = "Standardized coefficient estimates - leaf count")
plot_model(lfct.mod, show.values = T, type = "re")

# effect plots
plot_model(lfct.mod, type = "slope", show.data = T, colors = "blambus") +
      # geom_point(color = "black", alpha = .1) +
      # geom_smooth(method = "lm", color = "black", fill = "black") +
      # geom_smooth(method = "loess", color = NA) +
      theme_bw() +
      ylab("log (leaf count + 1)") +
      # theme(strip.text = element_blank()) +
      NULL

slc_vs_bay_plot <- plot_model(lfct.mod, type = "pred", terms = c("tot_bay.ln1"), show.data = F) +
      geom_point(data = oaks.data, aes(x = tot_bay.ln1, y = lfct),alpha = .1, size = 1) +
      theme_classic() +
      ylab("Symptomatic leaf count") +
      xlab("Bay laurel density") +
      ggtitle("Model fit") +
      NULL

slc_vs_hrs25_plot <- plot_model(lfct.mod, type = "pred", terms = c("hrs25"), show.data = F) +
      geom_point(data = oaks.data, 
                 aes(x = hrs25, y = lfct),
                 alpha = .1, size = 1) +
      theme_classic() +
      ylab("Symptomatic leaf count") +
      xlab("Hours >25 ºC") +
      ggtitle("Model fit") +
      NULL

# ggeffects::ggpredict(lfct.mod)

# slc_vs_bay_plot <- ggplot(oaks.data, aes(tot_bay.ln1, lfct)) +
#       geom_point(alpha = .7) +
#       geom_smooth(method = "lm", color = "black") +
#       xlab("log( Bay laurel count )") +
#       ylab("log( Symptomatic leaf count )") +
#       theme_classic()
# 
# slc_vs_hrs25_plot <- ggplot(oaks.data, aes(hrs25, lfct)) +
#       geom_point(alpha = .7) +
#       geom_smooth(method = "lm", color = "black") +
#       xlab("Number of hours >25 ºC") +
#       ylab("log( Symptomatic leaf count )") +
#       theme_classic()

cowplot::plot_grid(slc_vs_bay_plot, 
                   slc_vs_hrs25_plot,
                   labels = "auto",
                   nrow = 2)

# plot(resid(slc), main = "LMM of symptomatic leaf count residuals")
# qqnorm(resid(slc), main="Q-Q plot for conditional residuals - SLC")
# qqnorm(ranef(slc)$sample_year$`(Intercept)`, 
#        main="Q-Q plot for Sample Year")
# qqnorm(ranef(slc)$plotid$`(Intercept)`, 
#        main="Q-Q plot for plot ID")

# plot(slc@frame$umca_density_ha.log, resid(slc),
#      xlab = "bay laurel density",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. bay laurel density predictor")
# plot(slc@frame$hrs_abv25ds, resid(slc),
#      xlab = "# dry-season hours >25 C",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. dry-season hours >25 C predictor")
# plot(slc@frame$avg_hrs1422_wet_t_t1.log,
#      xlab = "avg. wet-hours 14-22C",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. wet-hours predictor")
# plot(slc@frame$H.2005, resid(slc),
#      xlab = "Shannon's diversity index",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. Shannon's diversity predictor")
# plot(slc@frame$twi15m.log.scl, resid(slc),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. wetness index predictor")

# ggplot(oaks.data, aes(tot_bay, tot_lfct)) +
#       geom_point() +
#       scale_x_log10() +
#       scale_y_log10()

```

```{r pairs plot of slc predictor variables, echo=F, eval=F}

pairs(na.omit(select(oakplots.sub, umca_density_ha.log, hrs_abv25ds, avg_hrs1422_wet_t_t1.log, H.2005, twi15m)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for symptomatic leaf count models")

```

## **Residual plots of average number of wet-hours model**

### Diagnostic plots
```{r wet.mod diganostic plots, echo=F}
# diagnostic plots
plot_model(wet.mod, type = "diag")
plot_model(wet.mod, type = "resid")

```

### Fitted model plots
```{r wet.mod fit plots}
# estimate plots
plot_model(wet.mod, show.values = T, type = "est", 
           title = "Coefficient estimates - wet hours")
plot_model(wet.mod, show.values = T, type = "std", 
           title = "Standardized coefficient estimates - wet hours")
plot_model(wet.mod, show.values = T, type = "re")

# effect plots
plot_model(wet.mod, type = "slope") +
      theme_bw() +
      # ylab("log (leaf count + 1)") +
      theme(strip.background = element_blank()) +
      NULL

# plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
# plot(wet_hours_1422@frame$umca_density_ha.log.scl, resid(wet_hours_1422),
#      xlab = "bay laurel density",
#      ylab = "model residuals",
#      main = "Wet-hours model residuals vs. bay laurel density predictor")
# plot(wet_hours_1422@frame$twi15m, resid(wet_hours_1422),
#      xlab = "bay laurel density",
#      ylab = "model residuals",
#      main = "Wet-hours model residuals vs. bay laurel density predictor")

```

## **Residual plots of dry-season temperature >25C model**

### Diagnostic plots
```{r hot.mod diagnostic plots, echo=F}

# diagnostic plots
plot_model(hot.mod, type = "diag")
plot_model(hot.mod, type = "resid")

```

### Fitted model plots
```{r hot.mod fit plots}
# estimate plots
plot_model(hot.mod, show.values = T, type = "est", 
           title = "Coefficient estimates - hours >25 ºC")
plot_model(hot.mod, show.values = T, type = "std", 
           title = "Standardized coefficient estimates - hours >25 ºC")
plot_model(hot.mod, show.values = T, type = "re")

# effect plots
plot_model(hot.mod, type = "slope", show.data = T) +
      # geom_point() +
      theme_classic() +
      ylab("Number of hours >25 ºC") +
      xlab("Topographic index") +
      theme(strip.background = element_blank(),
            strip.text = element_blank()) +
      NULL

# Fitted line with the observed data.
hot.mod.plot <- plot_model(hot.mod, type = "pred", terms = "twi15m", show.data = F) +
      geom_point(data = oaks.data, 
                         aes(x = twi15m, y = hrs25),
                         alpha = .1) +
      theme_classic() +
      ylab("Number of hours >25 ºC") +
      xlab("Topographic index") +
      theme(strip.background = element_blank(),
            strip.text = element_blank()) +
      NULL

(hot.mod.plot2 <- ggplot(oaks.data, 
                         aes(twi15m, hrs25,
                             # color = as.factor(year)
                             )
                         ) +
      geom_point(alpha = .1) +
      # geom_smooth(method = "lm", se = F) +
      geom_smooth(method = "lm", se = T, color = "black") +
      scale_color_viridis_d() +
      xlab("Topographic index") +
      ylab("Number of hours >25 ºC") +
      theme_classic() +
      NULL
      )

# plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
# plot(hrs_abv25@frame$twi15m, resid(hrs_abv25),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "Dry-season hours model residuals vs. wetness index predictor")
# plot(hrs_abv25@frame$umca_density_ha.log.scl, resid(hrs_abv25),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "Dry-season hours model residuals vs. wetness index predictor")

```

## **Plots of oak infection model**

### Diagnostic plots
```{r inf.mod diagnostic plots, echo=F}

plot_model(inf.mod, type = "diag")
plot_model(inf.mod, type = "resid")

```

### Fitted model plots
```{r inf.mod fit plots, echo=F}

cowplot::plot_grid(
      plot_model(inf.mod, type = "est", show.values = T, 
                 title = "Oak infection coefficient estimates",
                 vline.color = "gray80") +
            # geom_hline(yintercept = 1, color = "gray80",linetype = "dashed") +
            theme_classic(),
      plot_model(inf.mod, type = "std", show.values = T, 
                 title = "Oak infection std. coefficient estimates",
                 vline.color = "gray80"
                 ) +
            theme_classic(),
      nrow = 2
      )

inf.lfct.plot <- plot_model(inf.mod, type = "pred", terms = c("lfct"), show.data = T) +
      # geom_point() +
      theme_classic() +
      xlab("Symptomatic leaf count") +
      ylab("Infected oaks (%)") +
      ggtitle("") +
      NULL

# marginal effects of each significant predictor variable
cowplot::plot_grid(
inf.lfct.plot <- inf.lfct.plot +
      geom_point(data = oaks.data, 
                 aes(x = lfct, 
                     y = inf_oak_ct/(inf_oak_ct+uninf_oak_ct)
                     ),
                 alpha = .1, size = 1),
# inf.lfct.plot +
#       geom_point(data = filter(oaks.data, inf_oak_ct/(inf_oak_ct+uninf_oak_ct) <.5), 
#                  aes(x = lfct, 
#                      y = inf_oak_ct/(inf_oak_ct+uninf_oak_ct)
#                      ),
#                  alpha = .4, size = 1),

>>>>

---
title: "Pathogen Spillover - dissertation chapter"
author: "Whalen Dillon"
date: "July 18, 2016"
output:
  html_document:
    self_contained: no
    toc: yes
  pdf_document: default
  word_document: default
---

@@

[Conceptual Path Model](figures/conceptual-path-model-updated.png)

## Results

@@

```{r load data, echo=F, message=F}
library(knitr)
load("climate_path_model_data.RData")
library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
## Avoided this error in 1.1.2 by removing observations with NA values, which gives the same results as before, so maybe it wasn't properly dropping levels when extracting info from the models.
library(plyr); library(dplyr, verbose = F)
library(lme4, verbose = F); library(nlme, verbose = F)
# library(lmerTest, verbose = F)

@@

## make consolidated data frame only of variables in the path model
oakplots.model.data <- select(oakplots.sub, plotid, sample_year, oak.inf, tot_lfct, tot_lfct.log, hrs_abv25ds, avg_hrs1422_wet_t_t1, avg_hrs1422_wet_t_t1.log, umca_density_ha, umca_density_ha.log, H.2005, twi15m, twi15m.log#, twi15m.log.scl, tot_lfct.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, umca_density_ha.log.scl, shannons2005.scl
                              )

oakplots.model.data <- filter(oakplots.model.data, avg_hrs1422_wet_t_t1 !="NA", H.2005 != "NA", tot_lfct != "NA")
write.csv(oakplots.model.data, "analysis/data/oakplots-pathmodel-data.csv",
          row.names = F)
```

### Path model 
This is the list of models that make up the path model to be fit using the methods implemented in the 'piecewiseSEM' packag. The continuous variables in the models were standardized using the default settings of the 'scale' function in R, resulting in each having a mean of zero and standard deviation of one. Each variable is first centered by subtracting its mean and then scaled by dividing by its standard deviation.

@@

```{r hours >25 path model all oak plots unscaled, echo=F}
# slc_nlme <- lme(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005, random = ~1|plotid, data = oakplots.model.data)

model_list1 <- list(
      shannons <- lm(H.2005 ~ twi15m, data = filter(oakplots.model.data, sample_year == 2006)),

      umca_density <- lmer(umca_density_ha.log ~ twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),

@@

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ 
                                   # umca_density_ha.log + twi15m.log +
                                   twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      hrs_abv25 <- lmer(hrs_abv25ds ~ 
                              # umca_density_ha.log + 
                              twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
)

model_list1_refit <- list(
      shannons_refit <- lm(H.2005 ~ twi15m, data = filter(oakplots.model.data, sample_year == 2006)),

      umca_density_refit <- lmer(umca_density_ha.log ~ H.2005 + twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      slc_refit <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds +
                        #avg_hrs1422_wet_t_t1.log + H.2005 + 
                        (1|plotid) + (1|sample_year), 
                  data = oakplots.model.data),

      wet_hours_1422_refit <- lmer(avg_hrs1422_wet_t_t1.log ~ 
                                   # umca_density_ha.log + twi15m.log +
                                   twi15m +
                                   (1|plotid) + (1|sample_year), 
                             data = oakplots.model.data),

      hrs_abv25_refit <- lmer(hrs_abv25ds ~ 
                              # umca_density_ha.log + 
                              twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      oak_infection_refit <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
)
```

@@

### Examining the residual plots of each piece of the path model

**Residual plots of Shannon's Diversity Index model**
```{r shannons residuals plot, echo=F}
plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
plot(shannons$model$twi15m, resid(shannons),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Diversity model residuals vs. wetness index predictor")
```

**Residual plots of bay laurel density model**
```{r bay laurel density residuals plot, echo=F}
plot(umca_density, main = "LMM of bay laurel density residuals")
plot(umca_density@frame$twi15m, resid(umca_density),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Bay laurel density model residuals vs. wetness index predictor")
```

**Residual plots of symptomatic leaf count model**
```{r symptomatic leaf count residuals plot, echo=F}
plot(resid(slc), main = "LMM of symptomatic leaf count residuals")
# qqnorm(resid(slc), main="Q-Q plot for conditional residuals - SLC")
# qqnorm(ranef(slc)$sample_year$`(Intercept)`, 
#        main="Q-Q plot for Sample Year")
# qqnorm(ranef(slc)$plotid$`(Intercept)`, 
#        main="Q-Q plot for plot ID")

plot(slc@frame$umca_density_ha.log, resid(slc),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "SLC model residuals vs. bay laurel density predictor")
plot(slc@frame$hrs_abv25ds, resid(slc),
     xlab = "# dry-season hours >25 C",
     ylab = "model residuals",
     main = "SLC model residuals vs. dry-season hours >25 C predictor")
plot(slc@frame$avg_hrs1422_wet_t_t1.log,
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "SLC model residuals vs. wet-hours predictor")
plot(slc@frame$H.2005, resid(slc),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "SLC model residuals vs. Shannon's diversity predictor")
# plot(slc@frame$twi15m.log.scl, resid(slc),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. wetness index predictor")
```
```{r pairs plot of slc predictor variables, echo=F}
pairs(na.omit(select(oakplots.sub, umca_density_ha.log, hrs_abv25ds, avg_hrs1422_wet_t_t1.log, H.2005, twi15m)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for symptomatic leaf count models")
```

**Residual plots of average number of wet-hours model**
```{r wet-hours residuals plot, echo=F}
plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
# plot(wet_hours_1422@frame$umca_density_ha.log.scl, resid(wet_hours_1422),
#      xlab = "bay laurel density",
#      ylab = "model residuals",
#      main = "Wet-hours model residuals vs. bay laurel density predictor")
plot(wet_hours_1422@frame$twi15m, resid(wet_hours_1422),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "Wet-hours model residuals vs. bay laurel density predictor")
```

**Residual plots of dry-season temperature >25C model**
```{r dry-season hours residuals plot, echo=F}
plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
plot(hrs_abv25@frame$twi15m, resid(hrs_abv25),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Dry-season hours model residuals vs. wetness index predictor")
# plot(hrs_abv25@frame$umca_density_ha.log.scl, resid(hrs_abv25),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "Dry-season hours model residuals vs. wetness index predictor")
```

**Residual plots of oak infection model**
```{r oak infection residuals plot, echo=F}
plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
plot(oak_infection@frame$tot_lfct.log, resid(oak_infection),
     xlab = "symptomatic leaf count",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. symptomatic leaf count predictor")
plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log, resid(oak_infection),
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. wet-hours predictor")
plot(oak_infection@frame$H.2005, resid(oak_infection),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. diversity predictor")
```
```{r pairs plot of oak infection predictor variables, echo=F}
pairs(na.omit(select(oakplots.sub, tot_lfct.log, avg_hrs1422_wet_t_t1.log, H.2005)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for oak infection model")
```

### Fitting the path model
